TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-01-29 paper/draft.tex
improved figures
point to table

Another mode of analysis we can look at with \thecannon\ is how the derivative of the model changes with respect to given training parameters, which makes our model interpretable for discovering or verifying atomic or molecular lines with strong dependence on different physical parameters. The top two panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show two example spectra and model fits for two different temperatures (\ref{fig:demo_teff}), and two different temperatures (\ref{fig:demo_feh}), with atomic and molecular features identified by the abundance analysis of \citealt{Souto:2017}. The bottom panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show the derivative of flux with respect to temperature and metallicity at each pixel, taken at the median training values. In order to evaluate which spectral features show statistically significant change with respect to input label, we compute error of the derivative at each pixel using a jackknife statistic (with a 1-$\sigma$ level overplotted in red):
\begin{equation}
	\sigma_{\theta,m}^2 = \frac{N-1}{N} \sum^N_{n=1} (\theta_{/n}^m - \theta^m)^2 
\end{equation}
where $\sigma_{\theta,m}$ is the error at pixel \emph{m}, \emph{N} is the total number of stars in the sample, indexed by \emph{n}, $\theta$ is the coeffient vector trainined on all N sources, and $\theta_{/n}$ is the coefficient vector trained on \emph{N-1} sources excluding star \emph{n}.

\color{gcolor}{HOGG/ADAM: Do we want some note on what lines vary significantly with parameter? I could summarize the which kinds of lines show increasing or decreasing trends, but have no idea how to interpret from there.}\color{black}

@@

\begin{figure}[ht]
\begin{center}
\plotone{figures/self_and_validation_test_spt_snr.png}
\end{center}
\caption{Consistency test for the West-trained spectral type model; self-test (left), and leave-one-out cross validation (right).  Predictive accuracy, as computed by the scatter in cross validation is 0.9 subtypes.}\label{fig:west_validation}
\end{figure}

@@

\begin{deluxetable*}{|c|c|c|c|c|}
\tablecaption{Strong \teff\ correlated lines for all features with significance greater than $2\sigma_{\rm jackknife}$.}
\tablehead{Species & Wavelength & Derivative & $\sigma_{\rm jackknife}$ & Significance}
\startdata
Fe I & 15207.509 & -0.007 & 0.003 & -2.622 \\
Fe I & 15648.478 & -0.006 & 0.002 & -3.273 \\
Fe I & 15692.643 & -0.008 & 0.002 & -4.056 \\
FeH & 16107.127 & -0.002 & 0.001 & -2.78 \\
FeH & 16245.463 & -0.005 & 0.002 & -3.214 \\
FeH & 16271.519 & -0.002 & 0.001 & -2.013 \\
FeH & 16284.562 & -0.006 & 0.001 & -5.807 \\
FeH & 16377.066 & -0.005 & 0.001 & -4.554 \\
FeH & 16741.489 & -0.002 & 0.001 & -2.873 \\
FeH & 16812.415 & -0.004 & 0.001 & -4.241 \\
FeH & 16813.808 & -0.006 & 0.001 & -3.954 \\
FeH & 16922.405 & -0.003 & 0.001 & -3.867 \\
FeH & 16934.801 & -0.004 & 0.001 & -3.591 \\
OH & 15278.267 & 0.006 & 0.003 & 2.47 \\
OH & 15280.8 & 0.004 & 0.002 & 2.007 \\
OH & 15391.186 & 0.006 & 0.003 & 2.127 \\
OH & 15407.355 & 0.01 & 0.002 & 4.836 \\
OH & 15409.271 & 0.008 & 0.002 & 4.683 \\
OH & 15505.797 & 0.009 & 0.003 & 3.511 \\
OH & 15560.305 & 0.008 & 0.002 & 3.865 \\
OH & 15565.895 & 0.005 & 0.002 & 2.505 \\
OH & 15568.691 & 0.007 & 0.001 & 4.929 \\
OH & 15572.133 & 0.008 & 0.002 & 4.393 \\
OH & 16052.7 & 0.008 & 0.002 & 3.517 \\
OH & 16055.362 & 0.006 & 0.002 & 3.7 \\
OH & 16061.796 & 0.007 & 0.003 & 2.502 \\
OH & 16065.124 & 0.009 & 0.002 & 4.214 \\
OH & 16069.564 & 0.008 & 0.003 & 3.275 \\
OH & 16074.227 & 0.007 & 0.002 & 3.338 \\
OH & 16190.345 & 0.007 & 0.003 & 2.498 \\
OH & 16192.134 & 0.008 & 0.002 & 4.012 \\
OH & 16203.995 & 0.006 & 0.002 & 3.039 \\
OH & 16207.129 & 0.004 & 0.002 & 2.196 \\
OH & 16352.196 & 0.007 & 0.002 & 3.609 \\
OH & 16354.682 & 0.011 & 0.003 & 4.167 \\
OH & 16364.626 & 0.007 & 0.003 & 2.661 \\
OH & 16368.244 & 0.008 & 0.002 & 4.303 \\
OH & 16581.281 & 0.008 & 0.002 & 4.41 \\
OH & 16581.74 & 0.004 & 0.001 & 4.842 \\
OH & 16866.621 & 0.005 & 0.001 & 3.169 \\
OH & 16871.982 & 0.012 & 0.002 & 4.892 \\
OH & 16878.976 & 0.006 & 0.002 & 3.807 \\
OH & 16884.573 & 0.005 & 0.002 & 3.066 \\
OH & 16886.206 & 0.006 & 0.001 & 5.131 \\
OH & 16895.074 & 0.005 & 0.002 & 2.197 \\
Si I & 15960.044 & -0.009 & 0.004 & -2.523 \\
Si I & 16094.893 & -0.014 & 0.003 & -4.368 \\
Si I & 16680.77 & -0.01 & 0.003 & -3.475 \\
K I & 15162.824 & -0.006 & 0.002 & -2.294 \\
Ti I & 15715.641 & 0.006 & 0.002 & 2.894 \\
Ti I & 16634.973 & 0.003 & 0.001 & 4.224 \\
V I & 15923.924 & 0.003 & 0.001 & 3.6 \\
Mn I & 15159.054 & -0.006 & 0.002 & -3.123 \\
\enddata
\end{deluxetable*}

@@

\begin{deluxetable*}{|c|c|c|c|c|}
\tablecaption{Strong \feh\ correlated lines for all features with significance greater than $2\sigma_{\rm jackknife}$.}
\tablehead{Species & Wavelength & Derivative & $\sigma_{\rm jackknife}$ & Significance}
\startdata
Fe I & 15490.381 & -0.004 & 0.001 & -2.722 \\
Fe I & 15692.643 & -0.005 & 0.002 & -2.759 \\
Fe I & 16009.512 & -0.006 & 0.003 & -2.143 \\
FeH & 16114.027 & 0.004 & 0.001 & 2.673 \\
FeH & 16574.64 & 0.003 & 0.001 & 2.356 \\
FeH & 16694.372 & 0.003 & 0.001 & 2.193 \\
FeH & 16814.041 & 0.004 & 0.002 & 2.513 \\
OH & 16052.7 & 0.005 & 0.002 & 2.034 \\
OH & 16055.362 & 0.006 & 0.002 & 2.805 \\
OH & 16065.124 & 0.007 & 0.003 & 2.364 \\
OH & 16871.982 & 0.007 & 0.003 & 2.306 \\
OH & 16886.206 & 0.004 & 0.001 & 2.52 \\
Si I & 15960.044 & -0.007 & 0.003 & -2.112 \\
Si I & 16680.77 & -0.006 & 0.003 & -2.156 \\
K I & 15163.033 & 0.006 & 0.003 & 2.043 \\
V I & 15923.924 & 0.002 & 0.001 & 2.017 \\
Cr I & 15680.073 & -0.005 & 0.002 & -2.811 \\
Mn I & 15159.054 & -0.008 & 0.002 & -3.298 \\
Mn I & 15262.023 & -0.005 & 0.002 & -2.996 \\
\enddata
\end{deluxetable*}

>>>>

Another mode of analysis we can look at with \thecannon\ is how the derivative of the model changes with respect to given training parameters, which makes our model interpretable for discovering or verifying atomic or molecular lines with strong dependence on different physical parameters. The top two panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show two example spectra and model fits for two different temperatures (\ref{fig:demo_teff}), and two different temperatures (\ref{fig:demo_feh}), with atomic and molecular features identified by the abundance analysis of \citealt{Souto:2017}. The bottom panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show the derivative of flux with respect to temperature and metallicity at each pixel, taken at the median training values. In order to evaluate which spectral features show statistically significant change with respect to input label, we compute error of the derivative at each pixel using a jackknife statistic (with a 1-$\sigma$ level overplotted in red):
\begin{equation}
	\sigma_{\theta,m}^2 = \frac{N-1}{N} \sum^N_{n=1} (\theta_{/n}^m - \theta^m)^2 
\end{equation}
where $\sigma_{\theta,m}$ is the error at pixel \emph{m}, \emph{N} is the total number of stars in the sample, indexed by \emph{n}, $\theta$ is the coeffient vector trainined on all N sources, and $\theta_{/n}$ is the coefficient vector trained on \emph{N-1} sources excluding star \emph{n}. A summary of identified lines with derivative values greater than $2\sigma_{\rm jackkife}$ is given in Table \ref{table:teff_derivative} and Table \ref{table:feh_derivative}.

@@

\begin{figure}[ht]
\begin{center}
% \plotone{figures/self_and_validation_test_spt_snr.png}
\plottwo{figures/self_test_spt.png}{figures/crv_test_spt.png}
\end{center}
\caption{Consistency test for the West-trained spectral type model; self-test (left), and leave-one-out cross validation (right).  Predictive accuracy, as computed by the scatter in cross validation is 0.9 subtypes.}\label{fig:west_validation}
\end{figure}

@@

\begin{deluxetable*}{|c|c|c|c|c|}
\tablecaption{Strong \teff\ correlated lines for all features with significance greater than $2\sigma_{\rm jackknife}$.}
\tablehead{Species & Wavelength & Derivative & $\sigma_{\rm jackknife}$ & Significance}
\startdata
Fe I & 15207.509 & -0.007 & 0.003 & -2.622 \\
Fe I & 15648.478 & -0.006 & 0.002 & -3.273 \\
Fe I & 15692.643 & -0.008 & 0.002 & -4.056 \\
FeH & 16107.127 & -0.002 & 0.001 & -2.78 \\
FeH & 16245.463 & -0.005 & 0.002 & -3.214 \\
FeH & 16271.519 & -0.002 & 0.001 & -2.013 \\
FeH & 16284.562 & -0.006 & 0.001 & -5.807 \\
FeH & 16377.066 & -0.005 & 0.001 & -4.554 \\
FeH & 16741.489 & -0.002 & 0.001 & -2.873 \\
FeH & 16812.415 & -0.004 & 0.001 & -4.241 \\
FeH & 16813.808 & -0.006 & 0.001 & -3.954 \\
FeH & 16922.405 & -0.003 & 0.001 & -3.867 \\
FeH & 16934.801 & -0.004 & 0.001 & -3.591 \\
OH & 15278.267 & 0.006 & 0.003 & 2.47 \\
OH & 15280.8 & 0.004 & 0.002 & 2.007 \\
OH & 15391.186 & 0.006 & 0.003 & 2.127 \\
OH & 15407.355 & 0.01 & 0.002 & 4.836 \\
OH & 15409.271 & 0.008 & 0.002 & 4.683 \\
OH & 15505.797 & 0.009 & 0.003 & 3.511 \\
OH & 15560.305 & 0.008 & 0.002 & 3.865 \\
OH & 15565.895 & 0.005 & 0.002 & 2.505 \\
OH & 15568.691 & 0.007 & 0.001 & 4.929 \\
OH & 15572.133 & 0.008 & 0.002 & 4.393 \\
OH & 16052.7 & 0.008 & 0.002 & 3.517 \\
OH & 16055.362 & 0.006 & 0.002 & 3.7 \\
OH & 16061.796 & 0.007 & 0.003 & 2.502 \\
OH & 16065.124 & 0.009 & 0.002 & 4.214 \\
OH & 16069.564 & 0.008 & 0.003 & 3.275 \\
OH & 16074.227 & 0.007 & 0.002 & 3.338 \\
OH & 16190.345 & 0.007 & 0.003 & 2.498 \\
OH & 16192.134 & 0.008 & 0.002 & 4.012 \\
OH & 16203.995 & 0.006 & 0.002 & 3.039 \\
OH & 16207.129 & 0.004 & 0.002 & 2.196 \\
OH & 16352.196 & 0.007 & 0.002 & 3.609 \\
OH & 16354.682 & 0.011 & 0.003 & 4.167 \\
OH & 16364.626 & 0.007 & 0.003 & 2.661 \\
OH & 16368.244 & 0.008 & 0.002 & 4.303 \\
OH & 16581.281 & 0.008 & 0.002 & 4.41 \\
OH & 16581.74 & 0.004 & 0.001 & 4.842 \\
OH & 16866.621 & 0.005 & 0.001 & 3.169 \\
OH & 16871.982 & 0.012 & 0.002 & 4.892 \\
OH & 16878.976 & 0.006 & 0.002 & 3.807 \\
OH & 16884.573 & 0.005 & 0.002 & 3.066 \\
OH & 16886.206 & 0.006 & 0.001 & 5.131 \\
OH & 16895.074 & 0.005 & 0.002 & 2.197 \\
Si I & 15960.044 & -0.009 & 0.004 & -2.523 \\
Si I & 16094.893 & -0.014 & 0.003 & -4.368 \\
Si I & 16680.77 & -0.01 & 0.003 & -3.475 \\
K I & 15162.824 & -0.006 & 0.002 & -2.294 \\
Ti I & 15715.641 & 0.006 & 0.002 & 2.894 \\
Ti I & 16634.973 & 0.003 & 0.001 & 4.224 \\
V I & 15923.924 & 0.003 & 0.001 & 3.6 \\
Mn I & 15159.054 & -0.006 & 0.002 & -3.123 \\
\enddata
\end{deluxetable*} \label{table:teff_derivative}

@@

\begin{deluxetable*}{|c|c|c|c|c|}
\tablecaption{Strong \feh\ correlated lines for all features with significance greater than $2\sigma_{\rm jackknife}$.}
\tablehead{Species & Wavelength & Derivative & $\sigma_{\rm jackknife}$ & Significance}
\startdata
Fe I & 15490.381 & -0.004 & 0.001 & -2.722 \\
Fe I & 15692.643 & -0.005 & 0.002 & -2.759 \\
Fe I & 16009.512 & -0.006 & 0.003 & -2.143 \\
FeH & 16114.027 & 0.004 & 0.001 & 2.673 \\
FeH & 16574.64 & 0.003 & 0.001 & 2.356 \\
FeH & 16694.372 & 0.003 & 0.001 & 2.193 \\
FeH & 16814.041 & 0.004 & 0.002 & 2.513 \\
OH & 16052.7 & 0.005 & 0.002 & 2.034 \\
OH & 16055.362 & 0.006 & 0.002 & 2.805 \\
OH & 16065.124 & 0.007 & 0.003 & 2.364 \\
OH & 16871.982 & 0.007 & 0.003 & 2.306 \\
OH & 16886.206 & 0.004 & 0.001 & 2.52 \\
Si I & 15960.044 & -0.007 & 0.003 & -2.112 \\
Si I & 16680.77 & -0.006 & 0.003 & -2.156 \\
K I & 15163.033 & 0.006 & 0.003 & 2.043 \\
V I & 15923.924 & 0.002 & 0.001 & 2.017 \\
Cr I & 15680.073 & -0.005 & 0.002 & -2.811 \\
Mn I & 15159.054 & -0.008 & 0.002 & -3.298 \\
Mn I & 15262.023 & -0.005 & 0.002 & -2.996 \\
\enddata
\end{deluxetable*} \label{table:feh_derivative}
